Further Evidence for Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution

### 2021-12-20 src/ms.tex
paper edits

    \item The existence of the long-period pile-up limits the utility of gyrochronology for the hottest stars with convective envelopes, as stellar spin-down appears to stall on the ridge. For example, a Sun-like star may spend several Gyr evolving across the ridge. \textbf{Authors using rotation periods as means of identifying new members to coeval stellar associations should be cognizant of this pile-up for stars with \teff~$\gtrsim5800~K$ as it can mimic the the appearance of a coeval stellar association in the \teff--\prot plane.}

    \item An increasing number of open clusters are being discovered by searching for a clustering of rotation periods along a slow-rotator sequence. However, the long-period pile up discovered here can mimic a slow-rotator sequence in a small sample of unassociated stars with different ages and precisely measured temperatures and rotation periods. This is clearly demonstrated by the CKS sample in the right-hand panel of Figure~\ref{fig:surveys}. Taken out of context, this sample resembles a group of coeval stars with a slow-rotator sequence. The discovery of pile-ups in stellar rotation periods therefore has consequences for open cluster studies. When an overdensity or ridge is present in the rotation period distribution of a stellar population, care must be taken to ensure that it is not caused by WMB or core-envelope coupling before assuming that population is coeval and using the overdensity to age-date it.

>>>>

    \item The existence of the long-period pile-up limits the utility of gyrochronology for the hottest stars with convective envelopes, as stellar spin-down appears to stall on the pile-up. For example, a Sun-like star may spend several Gyr evolving across the ridge. Authors using gyrochronology as a means of age-dating a field dwarf star with \teff~$\gtrsim5800~K$ (and possibly cooler \teff as well) should take care to assess whether that star resides on the long-period edge, in which case the uncertainty on the age may be larger than current gyrochronology calibrations imply.

    \item An increasing number of open clusters are being discovered by searching for a clustering of rotation periods along a slow-rotator sequence in the \teff–\prot or color–\prot plane. However, the long-period pile up discovered here can mimic a slow-rotator sequence in a small sample of unassociated stars with different ages and precisely measured temperatures and rotation periods. This is clearly demonstrated by the CKS sample in the right-hand panel of Figure~\ref{fig:surveys}. Taken out of context, this sample resembles a group of coeval stars with a slow-rotator sequence. The discovery of pile-ups in stellar rotation periods therefore has consequences for open cluster studies. When an overdensity or ridge is present in the rotation period distribution of a stellar population, care must be taken to ensure that it is not caused by WMB or core-envelope coupling before assuming that population is coeval and using the overdensity to age-date it.
